---
title: "My Analysis of the BMI dataset"
author: "Melinda Higgins"
date: "March 11, 2016"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
```

## My Report on the BMI toy dataset

---

This is my first Rmarkdown report written as of `r Sys.time()`.

The dataset we will be using is a simple data example. We downloaded this example from Melinda Higgins' github website at [https://github.com/melindahiggins2000/CDCRworkshopCode](https://github.com/melindahiggins2000/CDCRworkshopCode).

```{r echo=FALSE}
data.csv <- read.csv(file="Dataset_01_comma.csv")
data.csv[18,"Height"] <- 5.6
data.csv$bmiPRE <- (data.csv$WeightPRE*703)/((data.csv$Height*12)**2)
data.csv$bmiPOST <- (data.csv$WeightPOST*703)/((data.csv$Height*12)**2)
data.csv$GenderFactor <- factor(data.csv$GenderCoded,
                                levels = c(1,2),
                                labels = c("Male","Female"))
data.csv <- within(data.csv,{
  bmiPREcat <- NA
  bmiPREcat[bmiPRE < 18.5] <- "underweight"
  bmiPREcat[bmiPRE >= 18.5 & bmiPRE < 25] <- "normal"
  bmiPREcat[bmiPRE >= 25 & bmiPRE < 30] <- "overweight"
  bmiPREcat[bmiPRE >= 30] <- "obese"
})
data.csv <- within(data.csv,{
  bmiPOSTcat <- NA
  bmiPOSTcat[bmiPOST < 18.5] <- "underweight"
  bmiPOSTcat[bmiPOST >= 18.5 & bmiPOST < 25] <- "normal"
  bmiPOSTcat[bmiPOST >= 25 & bmiPOST < 30] <- "overweight"
  bmiPOSTcat[bmiPOST >= 30] <- "obese"
})
data.csv$bmiPREcat.or <- factor(data.csv$bmiPREcat,
                                order = TRUE,
                                levels = c("underweight", 
                                           "normal",
                                           "overweight",
                                           "obese"))

data.csv$bmiPOSTcat.or <- factor(data.csv$bmiPOSTcat,
                                order = TRUE,
                                levels = c("underweight", 
                                           "normal",
                                           "overweight",
                                           "obese"))
```

### Data Table

Here is the data:

```{r echo=FALSE}
myvars <- c("Age", "bmiPRE", "bmiPREcat.or", 
            "bmiPOST", "bmiPOSTcat.or", 
            "GenderFactor")
data <- data.csv[,myvars]
names(data) <- c("Age",
                 "BMI PRE", "BMI PRE Category",
                 "BMI POST", "BMI POST Category",
                 "Gender")
kable(data, digits = 2)
```

### Plots of BMI Categories

Here is a barplot of the BMI categories at PRE

```{r echo=FALSE}
barplot(table(data.csv$bmiPREcat.or))
```

Here is a barplot of the BMI categories at PRE

```{r echo=FALSE}
barplot(table(data.csv$bmiPOSTcat.or))
```

### Comparisons of BMI PRE v POST

Here is a plot of the BMI PRE v POST. The linear model fit is shown with a **BLUE** line with 95% confidence intervals shown with grey shaded region and a **RED** line for the non-parametric _loess_ smoothed line.

```{r echo=FALSE}
data <- data.csv
p <- ggplot(data, aes(bmiPRE, bmiPOST)) + 
  geom_point() + 
  geom_smooth(method = "loess", colour = "red", se = FALSE) + 
  geom_smooth(method = "lm", colour = "blue") + 
  ggtitle("BMI PRE v POST: RED smoothed line, BLUE linear fit line")

p
```

### Linear Model Results

The best fit linear model results were:

```{r echo=FALSE}
fit1 <- lm(bmiPOST ~ bmiPRE, data=data.csv)
kable(as.data.frame(summary(fit1)$coefficients))
```

The ANOVA model fit for this model was:

```{r echo=FALSE}
kable(as.data.frame(anova(fit1)))
```

### Compare by Gender

Here is the BMI PRE v POST paneled by Gender.

```{r echo=FALSE}
data <- data.csv[!is.na(data.csv$GenderFactor),]
p <- ggplot(data, aes(bmiPRE, bmiPOST)) + 
  geom_point() + 
  geom_smooth(method = "loess", colour = "red", se = FALSE) + 
  geom_smooth(method = "lm", colour = "blue") + 
  facet_wrap(~GenderFactor) + 
  ggtitle("Panels for Gender, RED smoothed line, BLUE linear fit line")

p
```

